---
title: "**Title**"
subtitle: "**Supplmentary Material**"
author: "**Co-author names**"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 3
    toc-title: "**Table of Contents**"
    output-file: "index.html"
    theme: simplex
    embed-resources: true
    code-fold: true
    code-tools: true
    number-sections: true
    #bibliography: ./bib/ref.bib
    fontsize: "12"
    max-width: "10"
    code-overflow: wrap
crossref: 
  fig-title: Figure     # (default is "Figure")
  tbl-title: Table     # (default is "Table")
  title-delim: —     # (default is ":")
  fig-prefix: Fig.   # (default is "Figure")
  tbl-prefix: Tab.    # (default is "Table")
editor_options: 
  chunk_output_type: console
execute:
  warning: false
  message: false
  tidy: true
  #cache: true
---

Outstanding things and notes

- need to think what to do with levels less than 5 studies


## Setting up

### Loading packages

```{r}
pacman::p_load(tidyverse, # tidy family and related pacakges below
               kableExtra, 
               gridExtra, # may not use this
               pander,   # nice tables
               metafor,  # package for meta-analysis
               ape,      # phylogenetic analysis
               MuMIn,  # multi-model inference
               patchwork,   # putting ggplots together - you need to install via devtool
               here,         # making reading files easy
               orchaRd # plotting orchard plots

)

```

### Loading data: paper and tree data

```{r}
dat <- read.csv(here("data", "clean_data.csv"))

#dim(dat)
#head(dat)

tree <- read.tree(here("data", "species_tree.tre"))
```

### Custom functions

```{r}
# function to calculate effect sizes
# Zr - correlation
# there is always n

effect_size <- function(m1, m2, sd1, sd2, n1, n2, n, # 12 arguments
                        est , se, p_val, direction, method){

  if(method == "mean_method"){
  
    h <- n/n1 + n/n2
    p <- n1/n # prop for n1
    q <- n2/n # prop for n2
    s_pool <- sqrt( ((n1-1)*sd1^2 + (n2-1)*sd2^2) / (n - 2) )
    j <- 1 - (3 / (4*n - 9))
    d <- ((m2 - m1) / s_pool) * j
    r_pb <-  d / sqrt(d^2 + h)
    r_b <- r_pb*(sqrt(p*q)/dnorm(qnorm(p)))
    r <- r_b #r_b = r    
    
  }else if(method == "count_method"){
    
    h <- n/n1 + n/n2
    p <- n1/n # prop for n1
    q <- n2/n # prop for n2
    s1 <- sqrt(m1)
    s2 <- sqrt(m2)
    s_pool <- sqrt( ((n1-1)*s1^2 + (n2-1)*s2^2) / (n - 2) )
    j <- 1 - (3 / (4*n - 9))
    d <- ((m2 - m1) / s_pool) * j
    r_pb <-  d / sqrt(d^2 + h)
    r_b <- r_pb*(sqrt(p*q)/dnorm(qnorm(p)))
    r <- r_b #r_b = r     
    
  }else if(method == "percent_method1"){
    
    h <- n/n1 + n/n2
    p <- n1/n # prop for n1
    q <- n2/n # prop for n2
    m1 <- m1/100
    m2 <- m2/100
    s1 <- 1/sqrt(8)
    s2 <- 1/sqrt(8)
    m1 <- asin(sqrt(m1/100))
    m2 <- asin(sqrt(m2/100))
    s_pool <- sqrt( ((n1-1)*s1^2 + (n2-1)*s2^2) / (n - 2) )
    j <- 1 - (3 / (4*n - 9))
    d <- ((m2 - m1) / s_pool) * j
    r_pb <-  d / sqrt(d^2 + h)
    r_b <- r_pb*(sqrt(p*q)/dnorm(qnorm(p)))
    r <- r_b #r_b = r 
    
  }else if(method == "percent_method2"){
    
    h <- n/n1 + n/n2
    p <- n1/n # prop for n1
    q <- n2/n # prop for n2
    m1 <- m1/100
    m2 <- m2/100
    sd1 <- sd1/100
    sd2 <- sd2/100
    s1 <- 1/sqrt(sd1^2/(4*m1*(1-m1)))
    s2 <- 1/sqrt(sd2^2/(4*m2*(1-m2)))
    m1 <- asin(sqrt(m1/100))
    m2 <- asin(sqrt(m2/100))
    s_pool <- sqrt( ((n1-1)*s1^2 + (n2-1)*s2^2) / (n - 2) )
    j <- 1 - (3 / (4*n - 9))
    d <- ((m2 - m1) / s_pool) * j
    r_pb <-  d / sqrt(d^2 + h)
    r_b <- r_pb*(sqrt(p*q)/dnorm(qnorm(p)))
    r <- r_b #r_b = r 
    
  }else if(method == "proportion_method1"){
    
    h <- n/n1 + n/n2
    p <- n1/n # prop for n1
    q <- n2/n # prop for n2
    s1 <- 1/sqrt(8)
    s2 <- 1/sqrt(8)
    m1 <- asin(sqrt(m1))
    m2 <- asin(sqrt(m2))
    s_pool <- sqrt( ((n1-1)*s1^2 + (n2-1)*s2^2) / (n - 2) )
    j <- 1 - (3 / (4*n - 9))
    d <- ((m2 - m1) / s_pool) * j
    r_pb <-  d / sqrt(d^2 + h)
    r_b <- r_pb*(sqrt(p*q)/dnorm(qnorm(p)))
    r <- r_b #r_b = r 
    
  }else if(method == "proportion_method2"){
    
    h <- n/n1 + n/n2
    p <- n1/n # prop for n1
    q <- n2/n # prop for n2
    s1 <- 1/sqrt(sd1^2/(4*m1*(1-m1)))
    s2 <- 1/sqrt(sd2^2/(4*m2*(1-m2)))
    m1 <- asin(sqrt(m1/100))
    m2 <- asin(sqrt(m2/100))
    s_pool <- sqrt( ((n1-1)*s1^2 + (n2-1)*s2^2) / (n - 2) )
    j <- 1 - (3 / (4*n - 9))
    d <- ((m2 - m1) / s_pool) * j
    r_pb <-  d / sqrt(d^2 + h)
    r_b <- r_pb*(sqrt(p*q)/dnorm(qnorm(p)))
    r <- r_b #r_b = r 
    
  }else if(method == "t_method1"){
    

    p <- n1/n # prop for n1
    q <- n2/n # prop for n2
    r_pb <- est/sqrt(est^2 + n - 2)
    r_b <- r_pb*(sqrt(p*q)/dnorm(qnorm(p)))
    r <- r_b #r_b = r
    
  }else if(method == "t_method2"){
    
    n1 <- n/2
    n2 <- n/2
    p <- n1/n # prop for n1
    q <- n2/n # prop for n2
    r_pb <- est/sqrt(est^2 + n - 2)
    r_b <- r_pb*(sqrt(p*q)/dnorm(qnorm(p)))
    r <- r_b #r_b = r
    
  }else if(method == "F_method1"){
    
    p <- n1/n # prop for n1
    q <- n2/n # prop for n2
    r_pb <- sqrt(est)/sqrt(est + n -2)
    r_b <- r_pb*(sqrt(p*q)/dnorm(qnorm(p)))
    r_b = r_b*(direction)
    r <- r_b
  
  }else if(method == "F_method2"){
    
    n1 <- n/2
    n2 <- n/2
    p <- n1/n # prop for n1
    q <- n2/n # prop for n2
    r_pb <- sqrt(est)/sqrt(est + n -2)
    r_b <- r_pb*(sqrt(p*q)/dnorm(qnorm(p)))
    r_b = r_b*(direction)
    r <- r_b
    
  }else if(method == "p_method1"){
    
    p <- n1/n # prop for n1
    q <- n2/n # prop for n2
    t <- qt(1 - p_val, n - 2)
    r_pb <- t/sqrt(t^2 + n -2)
    r_b <- r_pb*(sqrt(p*q)/dnorm(qnorm(p)))
    r_b <- r_b*(direction)
    r <- r_b
    
  }else if(method == "p_method2"){
    
    n1 <- n/2
    n2 <- n/2
    p <- n1/n # prop for n1
    q <- n2/n # prop for n2
    t <- qt(1 - p_val, n - 2)
    r_pb <- t/sqrt(t^2 + n -2)
    r_b <- r_pb*(sqrt(p*q)/dnorm(qnorm(p)))
    r_b <- r_b*(direction)
    r <- r_b
    
  }else if(method == "correlation_method1"){
    
    r <- est
    
  }else if(method == "correlation_method2"){
    
    r <- 2*sin((pi/6)*est)
    
  }else if(method == "estimate_method1"){
    
    p <- n1/n # prop for n1
    q <- n2/n # prop for n2
    t <- est/se
    r_pb <- t/sqrt(t^2+ n -2)
    r_b <- r_pb*(sqrt(p*q)/dnorm(qnorm(p)))
    r <- r_b #r_b = r
  
  }else if(method == "estimate_method2"){
    
    n1 <- n/2
    n2 <- n/2
    p <- n1/n # prop for n1
    q <- n2/n # prop for n2
    t <- est/se
    r_pb <- t/sqrt(t^2+ n -2)
    r_b <- r_pb*(sqrt(p*q)/dnorm(qnorm(p)))
    r <- r_b #r_b = r
  
  } 
  
    if(r >= 1){
    # if over 1, we use 0.99 
    Zr <- atanh(0.99)
    }else if(r <= -1){
    Zr <- atanh(-0.99) # r = correlation
    } else {
    Zr <- atanh(r) # r = correlation
    } 
  
  VZr <- 1 /(n - 3)
  
  data.frame(ri = r, yi = Zr , vi = VZr)
  
}


#' Title: Contrast name generator
#'
#' @param name: a vector of character strings
cont_gen <- function(name) {
  combination <- combn(name, 2)
  name_dat <- t(combination)
  names <- paste(name_dat[, 1], name_dat[, 2], sep = "-")
  return(names)
}

#' @title get_pred1: intercept-less model
#' @description Function to get CIs (confidence intervals) and PIs (prediction intervals) from rma objects (metafor)
#' @param model: rma.mv object 
#' @param mod: the name of a moderator 
get_pred1 <- function (model, mod = " ") {
  name <- firstup(as.character(stringr::str_replace(row.names(model$beta), mod, "")))
  len <- length(name)
  
   if (len != 1) {
        newdata <- diag(len)
        pred <- metafor::predict.rma(model, 
                                     newmods = newdata,
                                     tau2.levels = 1:len)
    }
    else {
        pred <- metafor::predict.rma(model)
  }
  estimate <- pred$pred
  lowerCL <- pred$ci.lb
  upperCL <- pred$ci.ub 
  lowerPR <- pred$cr.lb
  upperPR <- pred$cr.ub 
  
  table <- tibble(name = factor(name, levels = name, labels = name), estimate = estimate,
                  lowerCL = lowerCL, upperCL = upperCL,
                  pval = model$pval,
                  lowerPR = lowerPR, upperPR = upperPR)
}

#' @title get_pred2: normal model
#' @description Function to get CIs (confidence intervals) and PIs (prediction intervals) from rma objects (metafor)
#' @param model: rma.mv object 
#' @param mod: the name of a moderator 
get_pred2 <- function (model, mod = " ") {
  name <- as.factor(str_replace(row.names(model$beta), 
                                paste0("relevel", "\\(", mod,", ref = name","\\)"),""))
  len <- length(name)
  
  if(len != 1){
  newdata <- diag(len)
  pred <- predict.rma(model, intercept = FALSE, newmods = newdata[ ,-1])
  }
  else {
    pred <- predict.rma(model)
  }
  estimate <- pred$pred
  lowerCL <- pred$ci.lb
  upperCL <- pred$ci.ub 
  lowerPR <- pred$cr.lb
  upperPR <- pred$cr.ub 
  
  table <- tibble(name = factor(name, levels = name, labels = name), estimate = estimate,
                  lowerCL = lowerCL, upperCL = upperCL,
                  pval = model$pval,
                  lowerPR = lowerPR, upperPR = upperPR)
}

#' @title mr_results
#' @description Function to put results of meta-regression and its contrasts
#' @param res1: data frame 1
#' @param res1: data frame 2
mr_results <- function(res1, res2) {
  restuls <-tibble(
    `Fixed effect` = c(as.character(res1$name), cont_gen(res1$name)),
    Estimate = c(res1$estimate, res2$estimate),
    `Lower CI [0.025]` = c(res1$lowerCL, res2$lowerCL),
    `Upper CI  [0.975]` = c(res1$upperCL, res2$upperCL),
    `P value` = c(res1$pval, res2$pval),
    `Lower PI [0.025]` = c(res1$lowerPR, res2$lowerPR),
    `Upper PI  [0.975]` = c(res1$upperPR, res2$upperPR),
  )
}


#' @title all_models
#' @description Function to take all possible models and get their results
#' @param model: intercept-less model
#' @param mod: the name of a moderator 

all_models <- function(model, mod = " ", type = "normal") {
  
  # getting the level names out
  level_names <- levels(factor(model$data[[mod]]))
  dat2 <- model$data
  mod <- mod


  run_rma1 <- function(name) {
      
    # variance covarinace matrix for sampling errors
    VCV1 <- vcalc(vi = dat$vi, cluster = dat$shared_group, rho = 0.5)
    VCV1 <- nearPD(VCV1)$mat
    
      
    rma.mv(yi, V = VCV1,
                   mods = ~relevel(dat2[[mod]], ref = name),
                     random = list(
                            ~ 1 | effectID,
                            ~ 1 | paperID,
                            ~ 1 | species_cleaned),
                          # ~ 1 | phylogeny),
             # R = list(phylogeny = cor_tree),
                     data = dat2,
                     test = "t",
                     sparse = TRUE,
                     control = list(optimizer = "Nelder-Mead"))
   }

    run_rma2 <- function(name) {
    
            # variance covarinace matrix for sampling errors
            VCVa <- vcalc(vi = dat2$vi_abs, cluster = dat$shared_group, rho = 0.5)
            VCVa <- nearPD(VCVa)$mat
               
               rma.mv(abs_yi2, V = VCVa,
               mods = ~relevel(dat2[[mod]], ref = name),
                      random = list(
                            ~ 1 | effectID,
                            ~ 1 | paperID,
                            ~ 1 | species_cleaned),
                     data = dat2,
                     test = "t",
                     sparse = TRUE,
                     control = list(optimizer = "Nelder-Mead"))
   }

# results of meta-regression including all contrast results; taking the last level out ([-length(level_names)])
# this does not work for hetero model?
if (type == "normal"){

    model_all <- purrr::map(level_names[-length(level_names)], run_rma1)

  } else {
  model_all <- purrr::map(level_names[-length(level_names)], run_rma2)
  }
  
  # getting estimates from intercept-less models (means for all the groups)
  res1 <- get_pred1(model, mod = mod)
  
  # getting estiamtes from all contrast models
  res2_pre <- purrr::map(model_all, ~ get_pred2(.x, mod = mod))
  
  # a list of the numbers to take out unnecessary contrasts
  contra_list <- Map(seq, from=1, to=1:(length(level_names) - 1))
  res2 <- purrr::map2_dfr(res2_pre, contra_list, ~.x[-(.y), ]) 
  # creating a table
  res_tab <- mr_results(res1, res2) %>% 
  kable("html",  digits = 3) %>%
  kable_styling("striped", position = "left") %>%
  scroll_box(width = "100%")
  
  # results
  res_tab

}


##########
# functions for absolute values


# folded mean
folded_mu <-function(mean, variance){
  mu <- mean
  sigma <- sqrt(variance)
  fold_mu <- sigma*sqrt(2/pi)*exp((-mu^2)/(2*sigma^2)) + mu*(1 - 2*pnorm(-mu/sigma))
  fold_mu
} 

# folded variance
folded_v <-function(mean, variance){
  mu <- mean
  sigma <- sqrt(variance)
  fold_mu <- sigma*sqrt(2/pi)*exp((-mu^2)/(2*sigma^2)) + mu*(1 - 2*pnorm(-mu/sigma))
  fold_se <- sqrt(mu^2 + sigma^2 - fold_mu^2)
  # adding se to make bigger mean
  fold_v <-fold_se^2
  fold_v
} 

```

### Preparing data: obtaining effect sizes etc. 


```{r}
# calcuating effect sizes
effect2 <- pmap_dfr(list(dat$mean_group_1, dat$mean_group_2, 
                         dat$variance_group_1, dat$variance_group_2, 
                         dat$n_group_1, dat$n_group_2, dat$n, 
                         dat$effect_size_value, dat$effect_size_variance, 
                         dat$effect_size_p_value_numeric, 
                         dat$direction_change, dat$function_needed), 
                    effect_size)                    


# merging two data frames
dat <- cbind(dat, effect2)

# adding absolute value (effect sizes) too
dat <- dat %>% mutate(
  abs_yi = abs(yi), # we use this one (conservative)
  abs_yi2 = folded_mu(yi, vi), # alternative way
  abs_vi = folded_v(yi, vi))


# dat$Zr <- unlist(effect1[1,])
# dat$VZr <- unlist(effect1[2,])



# renaming X to effectID
colnames(dat)[colnames(dat) == "X"] <- "effectID"

# creating the phylogeny column

dat$phylogeny <-  gsub(" ", "_", dat$species_cleaned)

# checking species names match between tree and dataset

# match(unique(dat$phylogeny), tree$tip.label)
# match(tree$tip.label, unique(dat$phylogeny))
# 
# intersect(unique(dat$phylogeny), tree$tip.label)
# setdiff(unique(dat$phylogeny), tree$tip.label)
# 
# match(unique(dat$phylogeny), tree$tip.label)
# sum(is.na(match(unique(dat$phylogeny), tree$tip.label)))

# creating a correlation matrix from a tree


# creating a VCV for sampling error variances 
# we assure shared groups will have r = 0.5
VCV <- vcalc(vi = dat$vi, cluster = dat$shared_group, rho = 0.5)
VCV <- nearPD(VCV)$mat


# creating VCV for absolute values
VCV_abs <- vcalc(vi = dat$abs_vi, cluster = dat$shared_group, rho = 0.5)
VCV_abs <- nearPD(VCV_abs)$mat


```

## Meta-analysis


## Phylogenetic mulitlevle model
::: {.panel-tabset}

## All random effects (5)

```{r}
#| eval: false
#| 
# takes a while to run
mod <- rma.mv(yi = yi, 
              V = VCV,
              mod = ~ 1,
              data = dat,
              random = list(
                            ~ 1 | effectID,
                            ~ 1 | paperID,
                            ~ 1 | species_cleaned,
                           ~ 1 | phylogeny),
              R= list(phylogeny = cor_tree),
              test = "t",
              sparse = TRUE)

# saving the runs
saveRDS(mod, here("Rdata", "mod.RDS"))
```

```{r}
# getting the save model
mod <- readRDS(here("Rdata", "mod.RDS"))

summary(mod)

# hardly any phlogenetic effects so taking it out
round(i2_ml(mod),2)   
```

## Reducted mulitlevle model

```{r}

# phylogeney accounts nothing (0) so we can take it out
mod1 <- rma.mv(yi = yi, 
               V = VCV,
              mod = ~ 1,
              data = dat, 
              random = list(
                ~ 1 | effectID,
                ~ 1 | paperID,
                ~ 1 | species_cleaned),
              test = "t",
              sparse = TRUE)

summary(mod1)

round(i2_ml(mod1),2)   

orchard_plot(mod1, xlab = "Effect Size: Zr", group = "paperID")

```


## Uni-moderator eta-regression models


!!! - model comparison we should use ML - !!!

## Reduced random effects (3)

```{r}
# phylogeney accounts nothing (0) so we can take it out
mod1 <- rma.mv(yi = yi, 
               V = VCV,
              mod = ~ 1,
              data = dat, 
              random = list(
                ~ 1 | effectID,
                ~ 1 | paperID,
                ~ 1 | species_cleaned),
              test = "t",
              sparse = TRUE)

summary(mod1)

round(i2_ml(mod1),2)   

orchard_plot(mod1, xlab = "Effect Size: Zr", group = "paperID")
```

:::

## Uni-moderator eta-regression models

::: {.panel-tabset}

## Sex

```{r}
mod_sex <- rma.mv(yi = yi, 
               V = VCV,
              mod = ~ sex - 1,
              data = dat, 
              random = list(
                ~ 1 | effectID,
                ~ 1 | paperID,
                ~ 1 | species_cleaned),
              test = "t",
              sparse = TRUE)
summary(mod_sex)

r2_ml(mod_sex)

orchard_plot(mod_sex, mod = "sex", xlab = "Effect Size: Zr", group = "paperID", branch.size = 4)


```

## Study type


```{r}
mod_type <- rma.mv(yi = yi,  
               V = VCV,
               mod = ~ study_type - 1,
               data = dat, 
               random = list(
                 ~ 1 | effectID,
                 ~ 1 | paperID,
                 ~ 1 | species_cleaned),
               test = "t",
               sparse = TRUE)
summary(mod_type)

r2_ml(mod_type)

orchard_plot(mod_type, mod = "study_type", xlab = "Effect Size: Zr", group = "paperID", branch.size = 4)


```


## Age classes

### Higher age grouping

```{r}
mod_age1 <- rma.mv(yi = yi,
               V = VCV,
               mod = ~ age_class_clean - 1,
               data = dat, 
               random = list(
                 ~ 1 | effectID,
                 ~ 1 | paperID,
                 ~ 1 | species_cleaned),
               test = "t",
               sparse = TRUE)
summary(mod_age1)
r2_ml(mod_age1)


orchard_plot(mod_age1, mod = "age_class", xlab = "Effect Size: Zr", group = "paperID", branch.size = 4)
```

### Finer age grouping

```{r}

# ageclass
mod_age2 <- rma.mv(yi = yi,
               V = VCV,
               mod = ~ age_class - 1,
               data = dat, 
               random = list(
                 ~ 1 | effectID,
                 ~ 1 | paperID,
                 ~ 1 | species_cleaned),
               test = "t",
               sparse = TRUE)
summary(mod_age2)
r2_ml(mod_age2)


orchard_plot(mod_age2, mod = "age_class", xlab = "Effect Size: Zr", group = "paperID", branch.size = 4)

```


## Adult or offspring fitness?

```{r}
mod_gen <- rma.mv(yi = yi, 
                V = VCV,
               mod = ~ whose_fitness - 1,
               data = dat, 
               random = list(
                 ~ 1 | effectID,
                 ~ 1 | paperID,
                 ~ 1 | species_cleaned),
               test = "t",
               sparse = TRUE)
summary(mod_gen)

orchard_plot(mod_gen, mod = "whose_fitness", xlab = "Effect Size: Zr", group = "paperID", branch.size = 4)

```

## Fitness metric

### Higher level

```{r}
mod_fit1 <- rma.mv(yi = yi, 
               V = VCV,
               mod = ~ fitness_higher_level - 1,
               data = dat, 
               random = list(
                 ~ 1 | effectID,
                 ~ 1 | paperID,
                 ~ 1 | species_cleaned),
               test = "t",
               sparse = TRUE)
summary(mod_fit1)
r2_ml(mod_fit1)

orchard_plot(mod_fit1, mod = "fitness_higher_level", xlab = "Effect Size: Zr", group = "paperID", branch.size = 4, angle = 45)
```

### Finer level

- need to think what to do with classes less than 5 studies

```{r}
mod_fit2 <- rma.mv(yi = yi, 
               V = VCV,
               mod = ~ fitness_metric_clean - 1,
               data = dat, 
               random = list(
                 ~ 1 | effectID,
                 ~ 1 | paperID,
                 ~ 1 | species_cleaned),
               test = "t",
               sparse = TRUE)
summary(mod_fit2)
r2_ml(mod_fit2)

orchard_plot(mod_fit2, mod = "fitness_metric_clean", xlab = "Effect Size: Zr", group = "paperID", branch.size = 4, angle = 45)
```

## Dispersal type

```{r}
mod_disp1 <- rma.mv(yi = yi, 
               V = VCV,
               mod = ~ dispersal_type - 1,
               data = dat, 
               random = list(
                 ~ 1 | effectID,
                 ~ 1 | paperID,
                 ~ 1 | species_cleaned),
               test = "t",
               sparse = TRUE)
summary(mod_disp1)
r2_ml(mod_disp1)

orchard_plot(mod_disp1, mod = "dispersal_type", xlab = "Effect Size: Zr", group = "paperID", branch.size = 4, angle = 45)
```

## Dispersal phase

```{r}
mod_disp2 <- rma.mv(yi = yi, V = vi,
               mod = ~ dispersal_phase - 1,
               data = dat, 
               random = list(
                 ~ 1 | effectID,
                 ~ 1 | paperID,
                 ~ 1 | species_cleaned),
               test = "t",
               sparse = TRUE)
summary(mod_disp2)
r2_ml(mod_disp2)

orchard_plot(mod_disp2, mod = "dispersal_phase", xlab = "Effect Size: Zr", group = "paperID", branch.size = 4, angle = 45)
```

## Study focus

### normal analysis

```{r}

mod12 <- rma.mv(yi = yi, 
                V = VCV,
                mod = ~ fitness_main_focus - 1,
                data = dat, 
                random = list(
                  ~ 1 | effectID,
                  ~ 1 | paperID,
                  ~ 1 | species_cleaned),
                # ~ 1 | phylogeny),
                # R= list(phylogeny = cor_tree),
                test = "t",
                sparse = TRUE)
summary(mod12)
r2_ml(mod12)
orchard_plot(mod12, mod = "fitness_main_focus", xlab = "Effect Size: Zr", group = "paperID", branch.size = 4, angle = 90)

```


```{r}
# hetero


```


## absolute value analysis

```{r}

mod11 <- rma.mv(yi = abs_yi2, 
                V = VCV_abs,
                mod = ~ fitness_main_focus - 1,
                data = dat, 
                random = list(
                  ~ 1 | effectID,
                  ~ 1 | paperID,
                  ~ 1 | species_cleaned),
                # ~ 1 | phylogeny),
                # R= list(phylogeny = cor_tree),
                test = "t",
                sparse = TRUE)
summary(mod11)
r2_ml(mod11)

```

```{r}

mod11a <- rma.mv(yi = abs_yi2, 
                V = VCV_abs,
                mod = ~ fitness_main_focus,
                data = dat, 
                random = list(
                  ~ 1 | effectID,
                  ~ 1 | paperID,
                  ~ 1 | species_cleaned),
                # ~ 1 | phylogeny),
                # R= list(phylogeny = cor_tree),
                test = "t",
                sparse = TRUE,
                method = "ML",
                control = list(optimizer = "Nelder-Mead"))
                
summary(mod11a)

# hetero
mod11b <- rma.mv(yi = abs_yi2, 
                 V = VCV_abs,
                 mod = ~ fitness_main_focus,
                 data = dat, 
                 random = list(
                   ~ fitness_main_focus | effectID,
                   ~ 1 | paperID,
                   ~ 1 | species_cleaned),
                 struct = "DIAG",
                 # ~ 1 | phylogeny),
                 # R= list(phylogeny = cor_tree),
                 method = "ML",
                 test = "t",
                 sparse = TRUE,
                 control = list(optimizer = "Nelder-Mead"))
summary(mod11b)

anova(mod11a, mod11b)


orchard_plot(mod11b, mod = "fitness_main_focus", xlab = "Effect Size: Zr", group = "paperID", branch.size = 4, angle = 90)

```

## Study design

```{r}
mod_design <- rma.mv(yi = yi, V = vi,
               mod = ~ study_design - 1,
               data = dat, 
               random = list(
                 ~ 1 | effectID,
                 ~ 1 | paperID,
                 ~ 1 | species_cleaned),
               test = "t",
               sparse = TRUE)
summary(mod_design)
r2_ml(mod_design)

orchard_plot(mod_design, mod = "study_design", xlab = "Effect Size: Zr", group = "paperID", branch.size = 4, angle = 45)
```


## Species class

```{r}
mod_class <- rma.mv(yi = yi, V = vi,
               mod = ~ species_class - 1,
               data = dat, 
               random = list(
                 ~ 1 | effectID,
                 ~ 1 | paperID,
                 ~ 1 | species_cleaned),
               test = "t",
               sparse = TRUE)
summary(mod_class)
r2_ml(mod_class)

orchard_plot(mod_class, mod = "species_class", xlab = "Effect Size: Zr", group = "paperID", branch.size = 4, angle = 45)
```


:::

## Multi-moderator-regression model


::: {.panel-tabset}

## Full model

```{r}
# use "ML" for model selection
#######################
# Mulit-variable models
#######################

dat_short$sln_duration <- scale(log(dat_short$duration_days))

# we need - to
mod_full <- rma.mv(yi = SMD, 
               V = VCVs, 
               random = list(~1|Spp , 
                             ~1 | RecNo,
                             ~1 | Obs_ID),
               mods =  ~ sln_duration*Type +
                         sln_duration*Treat_mod +
                         Sex,
               test = "t",
               method = "ML", # use "ML"
               sparse = TRUE,
               data = dat_short)
summary(mod_full)

round(r2_ml(mod_full)*100, 2)

```
 
## Visualization 
 
```{r}
orchard_plot(mod_full,
             mod = "Type",
             group = "RecNo", 
             xlab = "Standardised mean differnece (SMD)",
             branch.size = 3)

orchard_plot(mod_full,
             mod = "Treat_mod",
             group = "RecNo", 
             xlab = "Standardised mean differnece (SMD)",
             branch.size = 3)

# interaction plot of duration with trait types
int_type <- mod_results(mod_full, mod = "sln_duration", group = "RecNo", weights = "prop",
                                   by = "Type")

bubble_plot(int_type, group = "RecNo", mod = "sln_duration", xlab = "ln(duration in days) (z-transformed)",
                     legend.pos = "top.left", condition.nrow = 3, g = TRUE)

# interaction plot of duration with trait types
int_trt <- mod_results(mod_full, mod = "sln_duration", group = "RecNo", weights = "prop",
                        by = "Treat_mod")

bubble_plot(int_trt, group = "RecNo", mod = "sln_duration", xlab = "ln(duration in days)",
            legend.pos = "top.left", condition.nrow = 3, g = TRUE)
```
 
## Multi-model inference

```{r}
# mulit-model selection
candidates <- dredge(mod_full, trace = 2)

# displays delta AICc <2
candidates_aic2 <- subset(candidates, delta < 5) 
# model averaging
mr_averaged_aic2 <- summary(model.avg(candidates, delta < 5)) 

# relative importance of each predictor for all the models
importance <- sw(candidates)
```

 
::: 



## Publication bias

::: {.panel-tabset}


```{r}
####################
# publication bias
####################

funnel(mod1, 
       yaxis="seinv",
       # = "rstudent",
       xlab = "Standarized residuals",
       ylab = "Precision (inverse of SE)",
       ylim = c(0.001, 16),
       xlim = c(-10,15),
       digits=c(0,1)
)

# funnel(mod1, 
#        #yaxis="seinv",
#        # = "rstudent",
#        xlab = "Standarized residuals",
#        ylab = "Precision (inverse of SE)",
#        ylim = c(0.001, 16),
#        xlim = c(-10,15),
#        digits=c(0,1)
# )


```

## Egger regression: uni-moderator

```{r}
# write conditional statement here

# if each group n is avaiable - assume n/2
dat$effectN <- ifelse(is.na(dat$n_group_1), (dat$n/2)*2/dat$n,  
                      (dat$n_group_1 * dat$n_group_2) / (dat$n_group_1 + dat$n_group_2))
  
dat$sqeffectN <- sqrt(dat$effectN)

mod_egger <- rma.mv(yi = yi, 
                    V = VCV,
                mod = ~ sqeffectN,
                data = dat, 
                random = list(
                  ~ 1 | effectID,
                  ~ 1 | paperID,
                  ~ 1 | species_cleaned),
                # ~ 1 | phylogeny),
                # R= list(phylogeny = cor_tree),
                test = "t",
                sparse = TRUE)
summary(mod_egger)

round(r2_ml(mod_egger)*100, 2)

small <- bubble_plot(mod_egger,
                     mod = "sqeffectN",
                     group = "paperID",
                     xlab = "sqrt(Effective N)",
                     ylab = "Effect Size: Zr",
                     g = TRUE)

small
```

## Decline effect: uni-moderator

```{r}
# creating publication year from "reference"
dat$year <- with(dat, substr(reference, nchar(reference)-4, nchar(reference)))
dat$year <- as.integer(ifelse(dat$year == "2017a" | dat$year == "2017b", 2017, dat$year))
# decline effect
mod_dec <- rma.mv(yi = yi, V = vi,
                     mod = ~ year,
                     data = dat, 
                     random = list(
                       ~ 1 | effectID,
                       ~ 1 | paperID,
                       ~ 1 | species_cleaned),
                     # ~ 1 | phylogeny),
                     # R= list(phylogeny = cor_tree),
                     test = "t",
                     sparse = TRUE)
summary(mod_dec)

round(r2_ml(mod_dec)*100, 2)

decline <- bubble_plot(mod_dec,
                       mod = "year",
                       group = "paperID",
                       xlab = "Publication year",
                       ylab = "Effect Size: Zr",
                       g = TRUE)
decline 
```

## Combinmed analysis

```{r}
# All together

mod_comb <- rma.mv(yi = yi, 
                   V = VCV,
                   mod = ~ year + sqeffectN,
                   data = dat, 
                   random = list(
                     ~ 1 | effectID,
                     ~ 1 | paperID,
                     ~ 1 | species_cleaned),
                   # ~ 1 | phylogeny),
                   # R= list(phylogeny = cor_tree),
                   test = "t",
                   sparse = TRUE)
                   
summary(mod_comb)

round(r2_ml(mod_comb)*100, 2)

# small-study
bubble_plot(mod_egger,
            mod = "sqeffectN",
            group = "paperID",
            xlab = "sqrt(Effective N)",
            ylab = "Effect Size: Zr",
            g = TRUE)

# decline
bubble_plot(mod_comb,
            mod = "year",
            group = "paperID",
            xlab = "Publication year",
            ylab = "Effect Size: Zr",
            g = TRUE)

```

## Leave-1study-out (sensitivity analysis)

```{r}
#| eval: false

dat <- dat %>%
  mutate(leave_out = reference)

dat$leave_out <- as.factor(dat$leave_out)


LeaveOneOut_effectsize <- list()
for (i in 1:length(levels(dat$leave_out))) {
  temp_dat <- dat %>%
    filter(leave_out != levels(dat$leave_out)[i])
  
  VCV_leaveout <- vcalc(vi = temp_dat$vi, cluster = temp_dat$shared_group, rho = 0.5)
  
  LeaveOneOut_effectsize[[i]] <-  rma.mv(yi = yi,
                                         V = VCV_leaveout, 
                                         random = list(
                                           ~ 1 | effectID,
                                           ~ 1 | paperID,
                                           ~ 1 | species_cleaned),
                                         # ~ 1 | phylogeny),
                                         # R= list(phylogeny = cor_tree),
                                         test = "t",
                                         sparse = TRUE,
                                         data = temp_dat[temp_dat$leave_out != levels(temp_dat$leave_out)[i], ])
}

# writing function for extracting est, ci.lb, and ci.ub from all models
est.func <- function(model) {
  df <- data.frame(est = model$b, lower = model$ci.lb, upper = model$ci.ub)
  return(df)
}

# using dplyr to form data frame
MA_oneout <- lapply(LeaveOneOut_effectsize,function(x) est.func(x)) %>%
  bind_rows %>%
  mutate(left_out = levels(dat$leave_out))


# telling ggplot to stop reordering factors
MA_oneout$left_out <- as.factor(MA_oneout$left_out)
MA_oneout$left_out <- factor(MA_oneout$left_out, levels = MA_oneout$left_out)

# saving the runs
saveRDS(MA_oneout, here("Rdata", "MA_oneout.RDS"))


```

```{r}

MA_oneout <- readRDS(here("Rdata", "MA_oneout.RDS"))

# plotting
leaveoneout <- ggplot(MA_oneout) + geom_hline(yintercept = 0, lty = 2, lwd = 1) +

  geom_hline(yintercept = mod_ma2$ci.lb, lty = 3, lwd = 0.75, colour = "black") +
  geom_hline(yintercept = mod_ma2$b, lty = 1, lwd = 0.75, colour = "black") + 
  geom_hline(yintercept = mod_ma2$ci.ub,
             lty = 3, lwd = 0.75, colour = "black") + 
  geom_pointrange(aes(x = left_out, y = est,
                      ymin = lower, ymax = upper)) + 
  xlab("Study left out") + 
  ylab("Zr (effect size), 95% CI") +
  coord_flip() + 
  theme(panel.grid.minor = element_blank()) + theme_bw() + theme(panel.grid.major = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) + theme(axis.text.y = element_text(size = 6))

leaveoneout
```

 
::: 

## Figure gallary 



## R Session Informtion

```{r}
# pander for making it look nicer
sessionInfo() %>% pander()
```



