#code to creat a taxonomic tree
#last edtied Dec 12, 2023 by A. R. Martinig

options(scipen=999, dplyr.width = Inf, tibble.print_min = 50, repos='http://cran.rstudio.com/') #scipen forces outputs to not be in scientific notation #dplyr.width will show all columns for head() function and tibble.print_min sets how many rows are printed and repos sets the cran mirror

#load libraries

#to install ggtree and ggtreeExtra for the first time
#if (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
#BiocManager::install("ggtree")

#if (!requireNamespace("BiocManager", quietly=TRUE))
#    install.packages("BiocManager")
#BiocManager::install("ggtreeExtra")

packages=c("googlesheets4", "tidyverse", "rotl", "dplyr", "ape", "ggplot2", "ggnewscale", "ggtree", "ggtreeExtra") 

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
lapply(packages, library, character.only = TRUE)

select<-dplyr::select


# Specify the URL of your Google Sheet
url <- "https://docs.google.com/spreadsheets/d/1ZWQUCG5wuLIPJbhYfRW_rvFVb4qFiWtFm2e0BZ5WBkE/edit#gid=1046848762"

# Read the Google Sheet into a dataframe
Species_info <- read_sheet(url) %>%
	filter(!composite_variable=="Y", !obsID=="TBD") %>%
  select(c("reference", "paperID", "common_name", "species", "phylogeny")) 
  

#str(Species_info)
length(unique(Species_info$phylogeny)) #145
unique(Species_info$phylogeny) #needs cleaning
table(Species_info$phylogeny)

taxa <- tnrs_match_names(unique(Species_info$phylogeny)) 

length(unique(Species_info$phylogeny)) #145

taxa <- tnrs_match_names(unique(Species_info$phylogeny)) #runs fine
names(taxa)
synonyms(taxa) ##list of all synonyms
taxa$unique_name # main TOL names
table(taxa$approximate_match) ##0 approximate match
taxa %>% filter(approximate_match==TRUE)
table(taxa$flags) # flags 22 names with problems (hidden, hybrid, incertae_sedis_inherited, infraspecific, infraspecific) - will need fixing

mytree <- tol_induced_subtree(ott_ids = taxa$ott_id, label_format= "name") 
#this will fail

plot(mytree, show.tip.label = T, cex = 0.8, no.margin = TRUE)

is.binary(mytree) #if it returns FALSE that means you have several branches coming from the same node

mytree$tip.label

## Tree tip labels need some cleaning:
mytree$tip.label <- gsub("\\(.*", "", mytree$tip.label) #remove comments
mytree$tip.label <- gsub("_"," ", mytree$tip.label) #get rid of the underscores
mytree$tip.label <- trimws(mytree$tip.label) #getting rid of the trailing whitespace

mytree$tip.label

#this lets me check which species names don't match up
#you want to get 100% matching records between the two (your original labels and the labels in the tree)
sort(intersect(as.character(mytree$tip.label), unique(Species_info$phylogeny))) ## 136 names are matching -  got quite a few left to fix
sort(setdiff(unique(Species_info$phylogeny), as.character(mytree$tip.label))) ## names from species column not matching with tip label
sort(setdiff(as.character(mytree$tip.label), unique(Species_info$phylogeny))) ## names from tip.label not matching in animal column

#need to fix any of the substitutions AND labels that are matched, but incorrectly

## Reverse substitutions of species names in the tree tip labels:
#mytree$tip.label[mytree$tip.label == "Vulpes macrotis"] <- "Vulpes macrotis mutica" 
mytree$tip.label[mytree$tip.label == "Macaca mulatta"] <- "Macaca mulatta tcheliensis" 
	


#check matching records again
sort(intersect(as.character(mytree$tip.label), unique(Species_info$species))) ## 145 names are matching - all fixed 

plot(mytree, show.tip.label = T, cex = 0.8, no.margin = TRUE)
str(mytree) #288 tips

write.tree(mytree, file = "~/Documents/Files/Post-docs/UNSW 2022-2024/Aim 1/Fitness-and-dispersal-MA/data/species_tree.tre") #save the tree 
 mytree <- read.tree(file = "~/Documents/Files/Post-docs/UNSW 2022-2024/Aim 1/Fitness-and-dispersal-MA/data/species_tree.tre") #if you need to read the tree

write.csv(Species_info, "~/Documents/Files/Post-docs/UNSW 2022-2024/Aim 1/Fitness-and-dispersal-MA/data/species_tree.csv")

head(Species_info)

#these tip labels are only to be used to relabel for the final plot, not for the data that is being shared

mytree$tip.label[mytree$tip.label == "Vulpes macrotis"] <- "Vulpes macrotis mutica" #kit fox
mytree$tip.label[mytree$tip.label == "Junco hyemalis"] <- "Passeriformes"
mytree$tip.label[mytree$tip.label == "Canis lupus"] <- "Canis lupus & Canis lupus arabs"
mytree$tip.label[mytree$tip.label == "Lagopus lagopus"] <- "Lagopus lagopus & Lagopus lagopus scoticus"
mytree$tip.label[mytree$tip.label == "Panthera tigris"] <- "Panthera tigris & Panthera tigris tigris"
mytree$tip.label[mytree$tip.label == "Calidris alpina"] <- "Calidris alpina schinzii" 

```

```{r plot tree}

# count of papers per species name:
sample_data <- count(Species_info,species)
 
sample_data$species %in% mytree$tip.label
mytree$tip.label %in% sample_data$species
 
 ggtree(mytree, layout="circular") +                    
   geom_tiplab( ) 
 
 
ggtree(mytree, layout="circular") +                    
   geom_tiplab(                             # adds name of sample to tip of its branch 
     color = 'black',                       
     offset = 1,
     size = 1,
     geom = "text",
     align = TRUE) 
 
ggtree(mytree, layout = "circular", branch.length = 'none') %<+% sample_data + # %<+% adds dataframe with sample data to tree 
   new_scale_color() +                             # allows to add an additional color scheme for another variable
     geom_tippoint(
       mapping = aes(color = n),  
       size = 1) +
   scale_colour_gradient(low = "yellow", high = "red") +
   geom_tiplab(                             # adds name of sample to tip of its branch 
     color = 'black',                       
     offset = 1,
     size = 1,
     geom = "text",
     align = TRUE) 
   
 
ggtree(mytree, layout="circular", branch.length = "none") # simple circular tree with all tips aligned
ggtree(mytree, layout="fan", open.angle=10, yscale = "none")
 
ggsave("species_tree_circular_v0.pdf", width = 12, height = 10, scale = 10)
 
p
```

```{r plot tree}

sample_data <- count(Species_info, common_name)
sample_data <- as.data.frame(sample_data)
sample_data2 <-mutate(sample_data, tip.label = common_name)
mytree$tip.label <- strip_ott_ids(mytree$tip.label, remove_underscores = TRUE)
 summary(sample_data2$tip.label)
 str(sample_data2)
 
 sample_data3<-left_join(sample_data2, Species_info, by= "common_name")
 sample_data4<-distinct(sample_data3, common_name, .keep_all=TRUE) %>%
 	mutate_if(is.character, as.factor)
 str(sample_data4)

#logging n
sample_data4$n_logged <- log(sample_data4$n)

#############

#############


#losia's code is below

cols <- c("yellow", "violet", "turquoise", "tomato", "thistle", "springgreen", "navy", "orange", "lightgoldenrod", "lightpink", "khaki", "lightslateblue", "red", "seagreen", "plum", "hotpink4", "moccasin", "olivedrab")

tree2 <- ggtree(mytree, layout = "circular", lwd = 0.75) %<+% sample_data4 + aes(col = species)+ theme(legend.position = "bottom") + scale_colour_manual(values = cols, (title = "Broad taxa")) 
tree2


tree3 <- tree2 + new_scale_fill() + geom_fruit(geom = geom_bar, mapping = aes(x = log(n), col = "gray30"), stat = "identity", col = "gray30", orientation = "y", axis.params = list(axis = "x",text.angle = -45, hjust = 0, text.size = 3), grid.params = list(alpha = 0.35),offset = 0.085, pwidth = 0.55, alpha = 0.8) + guides(fill ="none") 

tree3

count(sample_data4, common_name)
count(sample_data4, species)

## will add silhouettes to most common species
##need to manually change log axis back to original numbers

```
#next step is to add the sample size first
#make the text italics
#can add four main silhouttes - mammals, birds, fish, insects (four colours)

#can create a new column to pull the higher levels for each species -> need to assign to higher group
#there is a website with a database for silhouettes -> can get them for free online
#https://www.google.ca/search?source=hp&ei=y822X-3WI9Tk-gSz2o_QBw&q=free+animal+silhouettes+linked+for+R&btnK=Google+Search&oq=game+store+calgary&gs_lcp=CgZwc3ktYWIQAzIFCAAQyQMyAggAMgIIADIGCAAQFhAeMgYIABAWEB4yBggAEBYQHjIGCAAQFhAeMgYIABAWEB4yBggAEBYQHjIGCAAQFhAeOggIABCxAxCDAToFCAAQsQM6CwguELEDEMcBEKMCOgUIABCSAzoFCC4QsQM6CggAELEDEIMBEAo6AgguOggILhDHARCjAjoICAAQsQMQyQM6DgguELEDEIMBEMcBEKMCOggILhCxAxCTAjoQCC4QsQMQgwEQxwEQowIQCjoECAAQCjoICC4QxwEQrwFKBQgHEgExSgUICRIBMUoHCAoSAzEwMFDcF1jHJGDmJmgAcAB4AIABrwGIAesNkgEEMTMuNZgBAKABAaoBB2d3cy13aXo&sclient=psy-ab&ved=0ahUKEwit6L7tsY_tAhVUsp4KHTPtA3oQ4dUDCAk&uact=5#:~:text=PhyloPic,www.phylopic.org
# r package: https://cran.r-project.org/web/packages/rphylopic/index.html